Hybrid Threads for the Parrot Virtual Machine

### 2012-02-25 Paper.tex
Thread creation

>>>>

Each thread is represented by an instance of the ParrotInterpreter PMC. These interpreters are kept in the \textit{threads\_array} defined in \textit{src/thread.c}. It's a simple C array instead of one of Parrot's dynamic arrays which are implemented as PMCs because though ParrotInterpreter is implemented as a PMC itself, it's not fully garbage collected because of the bootstrapping issues this would create.

\textit{src/thread.c} also contains thread management functions all prefixed by \textit{Parrot\_thread\_}. Threads are created when \textit{Parrot\_cx\_schedule\_task} determines that all existing threads are busy and new threads may yet be started. It uses \textit{Parrot\_thread\_create} which creates a new interpreter by cloning the current one but with a cleared out \textit{thread\_data} structure. It also creates the \textit{thread\_data->notifierfd} pipe used to notify threads of expired alerts.

The task is then scheduled with the new interpreter's scheduler using \textit{Parrot\_thread\_schedule\_task}. This function uses \textit{Parrot\_thread\_create\_local\_task} to create a corresponding task on the new interpreter and then simply pushes it onto the scheduler. Since each interpreter has its own garbage collector, all objects used on the interpreter must origin from this GC's memory pools. That's why \textit{Parrot\_thread\_create\_local\_task} creates a new \textit{local\_task}. Proxies are created for the task's \textit{code} and \textit{data} attributes. The task object of the originating interpreter and the new \textit{local\_task} are linked using their \textit{partner} attributes which just contains a pointer to the other object. This is the point where for all variables in the original task's \textit{shared} array proxies are created an put into the \textit{shared} array of the \textit{local\_task}. \textit{Parrot\_thread\_create\_proxy} is used for this purpose.

\textit{Parrot\_thread\_insert\_thread} is used to put the new interpreter into the \textit{threads\_array}.

\textit{Parrot\_thread\_run} is the function where the actual operating system thread is created. It uses macros defined in the \textit{include/parrot/thr\_*.h} files which act as an OS abstraction layer.

\textit{Parrot\_thread\_outer\_runloop} is used as the thread's main function. It's very similiar to \textit{Parrot\_cx\_outer\_runloop}. The most important difference is the way it waits for pending alarms in case no tasks are to be run. While \textit{Parrot\_cx\_outer\_runloop} waits for an alarm signal to arrive, \textit{Parrot\_thread\_outer\_runloop} waits for a dummy write to be written to its \textit{notifierfd} pipe by \textit{Parrot\_cx\_run\_scheduler}. In addition it's holding the \textit{interp->thread\_data->interp\_lock} lock while executing a task. This is the lock which \textit{Parrot\_cx\_schedule\_task} is taking during its call to \textit{Parrot\_thread\_schedule\_task} because the latter needs to create new objects local to the thread. As this could cause garbage collection to be triggered, the thread itself should not be executing code at the same time.
